---
title: "Example of using `rss_varbvsr_naive`"
author: "Xiang Zhu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rssr usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this vignette, we show how to use `rssr::rss_varbvsr_naive` on a simulated dataset ([details](https://github.com/stephenslab/rss/wiki/Example-2#details)), and then compare this `R` function with its `Matlab` counterpart, [`rss_varbvsr.m`](https://github.com/stephenslab/rss/blob/master/src_vb/rss_varbvsr.m).

Before running this example, please make sure that [`rssr`](https://github.com/stephenslab/rssr) package is installed. If you have trouble installing this package, please open an [issue](https://github.com/stephenslab/rssr/issues) and provide as many details as possible.

```{r, eval=FALSE}
devtools::install_github("stephenslab/rssr",ref="v0.1.0-alpha")
```

## Use `rss_varbvsr_naive`

### Step 1: load summary data and correlation matrix

The input data for fitting an `RSS` model consist of i) a vector of marginal effect estimates (`betahat`); ii) a vector of standard errors of these marginal estimates (`se`) and iii) a correlation matrix (`R`) among all covariates. To make the model fitting results reproducible, here we also provide initial values of variational parameters (`alpha_start` and `mu_start`).

```{r, message=FALSE}
suppressPackageStartupMessages(library(rssr))
suppressPackageStartupMessages(library(Matrix))

# marginal effect estimates 
data("betahat")
betahat <- c(betahat)

# standard errors of marginal effect estimates
data("se")
se <- c(se)

# shrinkage estimate of correlation matrix
# note that this matrix is sparse: look at `is(R)`
data("R_shrink")
R <- R_shrink

# initial values of variational parameters 
data("alpha_test")
data("mu_test")

# convert row vectors to column vectors
mu_start <- t(t(mu_test))
alpha_start <- t(t(alpha_test))
```

We also compute a matrix `SiRiS`, defined as `SiRiS = diag(1/se) %*% as.matrix(R) %*% diag(1/se)`, before we start the variational update. 

```{r}
# compute the matrix SiRiS
SiRiS_f <- as.matrix(rssr::SiRSi(R, 1/se))
# note that SiRiS can be also computed as follows:
# SiRiS_f <- diag(1/se) %*% as.matrix(R) %*% diag(1/se)

# convert SiRiS to a sparse matrix
SiRiS <- as(SiRiS_f,"dgCMatrix")

# compute the initial value of SiRiSr
SiRiSr_start <- c(SiRiS_f %*% (alpha_start * mu_start))
```

### Step 2: fit the RSS-BVSR model via variational approximation

Here we fix the values of two hyper-parameters: `sigma_beta=1` and `logodds=-4.6`.

```{r}
r_results <- rssr::rss_varbvsr_naive_sp(SiRiS, sigma_beta=1, logodds=-4.6, betahat, se, alpha_start, mu_start, SiRiSr_start,itermax=100,verbose = F,lnz_tol = F, tolerance=1e-4)

# extract certain quantities for further comparisons
r_iter <- r_results$iter
r_maxerr <- r_results$max_err
r_lnZ <- r_results$lnZ
r_alpha <- r_results$alpha
r_mu <- r_results$mu
```

## Use `rss_varbvs.m`

Now we run the same analyses on the same data in `Matlab`, using `rss_varbvsr.m`. Please note that this part **cannot** be completed within `R`. We provide the result file for this part, so you can skip this part. If you would like to run this part in your own machine, please make sure the `rss_varbvsr.m` is installed first ([details](https://github.com/stephenslab/rss/wiki/RSS-via-VB)).  

```matlab
% run this example under the folder "examples" 
addpath(genpath('../src_vb'));
addpath('/home/xiangzhu/varbvs-master/varbvs-MATLAB/');

% load summary-level data
example_data = matfile('example2.mat');

betahat = example_data.betahat;   % single-SNP effect size estimates
se      = example_data.se; 	      % standard error of betahat

% load SNP correlation (a.k.a. LD) matrix
genotype_data = matfile('genotype2.mat');

R = genotype_data.shrink_R;       % shrinkage estimate of correlation matrix

% compute SiRiS matrix
p     = length(betahat);
Si    = 1 ./ se;
SiRiS = repmat(Si, 1, p) .* R .* repmat(Si', p, 1);
SiRiS = sparse(SiRiS);

% specify hyper-parameters
sigb    = 1;
logodds = -4.6;

% specify initial values for variational parameters
start_value = matfile('example2_start.mat');
alpha0      = start_value.alpha_start;
mu0         = start_value.mu_start;
options     = struct('alpha',alpha0,'mu',mu0,'verbose',false);

% run rss-varbvsr
[logw,alpha,mu,s,info] = rss_varbvsr(betahat,se,SiRiS,sigb,logodds,options);

% save results
save('example2_vb_results.mat','logw','alpha','mu','s','info');
```

## Compare results from `rss_varbvsr_naive` and `rss_varbvs.m`

Because `rss_varbvsr_naive` and `rss_varbvs.m` are written based on the **same** algorithm, their results should be the **same** if the input data and initial values are the **same**. Here we confirm this claim.  

If you do not run the `Matlab` script above to generate results from `rss_varbvsr.m`, please download the our result file [`example2_vb_results.mat`](https://guchicao.box.com/v/example2) to an directory (e.g. `/tmp`).

```{r,eval=F}
matfile_name <- system.file("example2_vb_results.mat",package="rssr")
m_results <- R.matlab::readMat(matfile_name)

# number of iterations
m_iter <- c(m_results$info[[1]])

# maximum relative difference between the parameters at the last two iterations
m_maxerr <- c(m_results$info[[2]])

# variational lower bound at the last iteration
m_lnZ <- c(m_results$logw)

# variational parameter estimates
m_alpha <- c(m_results$alpha)
m_mu <- c(m_results$mu)
```

First, we can see that both implementations yield the same number of iterations (`iter`), the same maximum relative difference between the parameters at the last two iterations (`maxerr`), and the same variational lower bound at the last iteration (`lnZ`).  

```{r,eval=F, echo=FALSE}
tab.mat <- matrix(data=0, nrow=3, ncol=2)
tab.mat[, 1] <- c(r_iter, log10(r_maxerr), r_lnZ)
tab.mat[, 2] <- c(m_iter, log10(m_maxerr), m_lnZ)

rownames(tab.mat) <- c("`iter`", "`log10(maxerr)`", "`lnZ`")
colnames(tab.mat) <- c("`rss_varbvsr_naive`", "`rss_varbvsr.m`")

suppressPackageStartupMessages(library(knitr))
knitr::kable(tab.mat)
```

In addition, the variational parameter estimates (`alpha` and `mu`) are identical.

```{r,eval=F, echo=FALSE, fig.width=8, fig.height=4}
par(mfrow=c(1,2))

plot(r_mu, m_mu, pch=20, xlab="rss_varbvsr_naive", ylab="rss_varbvsr.m", main="mu")
abline(0,1,col="red")

plot(r_alpha, m_alpha, pch=20, xlab="rss_varbvsr_naive", ylab="rss_varbvsr.m", main="alpha")
abline(0,1,col="red")
```

## Session information

```{r, echo=FALSE}
sessionInfo()
```
